---
title: "statinf_cp1_joannanw"
author: "joannanw"
date: "Sunday, March 22, 2015"
output: word_document
---
# This project investigates the exponential distribution in R and compare it with the Central Limit Theorem. The distribution is made up of the average of 40 exponentials ran 1000 times. 


# Simulations
Lambda is set to 0.2 for this simulation. The theoretical mean of distribution is first calculated. Following that, a sample simulation of the average of 40 exponentials is ran 1000 times. 

```{r}
lambda <- 0.2
theoreticalMean <- 1/lambda
```

The theoretical Mean is `r theoreticalMean`.

```{r}
set.seed(1)
sampleSize <- 40
sampleRun <- 1000
data <- NULL
for (i in 1 : sampleRun) data <- c(data, mean(rexp(sampleSize,lambda)))
dataMean <- mean(data)
```
The sample mean is `r dataMean`.

The theoretical and sample means are very close to each other. This shows that for the large simulation of a distribution, the average centers to its  theoretical mean.

```{r}
theoreticalVar <- (1/lambda)^2/sampleSize
dataVar <- var(data)
```

The theoretical variance is `r theoreticalVar` and the sample variance is `r dataVar`. The variability of the sample is close to the theoretical value. 

# Results

Now let's show the distribution of the sample.

```{r}
library(ggplot2)
sampleData <- data.frame(Type = "Sample", Data = data, stringsAsFactors = "FALSE")
bothMeans <- data.frame(Type = c("Sample", "Theoretical"), Mean = c(dataMean, theoreticalMean))
ggplot(sampleData, aes(x = Data)) + geom_histogram(alpha = 0.6, fill = "pink") + ggtitle("Sample Means for Exponential Distribution with Lambda = 0.2") + scale_y_continuous("Number of Observations") + geom_vline(data = bothMeans, aes(xintercept = Mean, color = Type, linetype = Type), show_guide = TRUE)
qqnorm(data, col = "red")
qqline(data, col = 1)
```

The distribution of the sample follows the Central Limit Theorem (CLT) where the arithmetic average and variance of the sample are well-defined and converges to a normal distribution.This is further shown by the Q-Q plot that the sample data lies very closely to the normal distribution. 





